SLT_paper

### 2020-08-05 SLT_paper.Rmd
first full version of analyses, knits

---
title: "SLT_paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)

@@

```{r}
read.csv("first_class_lisa.csv")->zoon
summary(zoon)

read.csv("Metadata_batch2_lisa.csv",sep="\t")->md
summary(md)
#? correct errors in md? onset           offset       chunk_pos

read.csv("result_final_lisa.csv")->lab
summary(lab)

data_all <- read.csv("final_classifications_PU_zoon.csv")

@@

data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",
                                                             "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))

# create lab column with easier to read correspondance
data_all$lab<-as.character(data_all$Major_Choice)
data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))

## correspondence
i) quantifying the correspondence between expert human annotations and Zooniverse annotations %(confusion matrix) 
and

@@

@@

@@

@@

## What is said about infants

ii) examining the correspondences between descriptors at the child level across the two annotation settings (e.g. CR-AGE correlations).

```{r by child}

@@

zoon_sums = data_all %>% group_by(ChildID,Zoon_classif) %>% tally()
colnames(zoon_sums)<-c("ChildID",      "Zoon_classif", "n_zoo")
lab_sums = data_all %>% group_by(ChildID,lab) %>% tally()
zoon_sums$id=paste(zoon_sums$ChildID,zoon_sums$Zoon_classif)
lab_sums$id=paste(lab_sums$ChildID,lab_sums$lab)
sums=merge(zoon_sums,lab_sums[,c("lab" ,  "n", "id")],by="id",all=T)

@@

# Canonical Rate analysis by Age & ChildID

```{r CR by Age}
df.sub <- subset(data_all, Zoon_classif %in% c("Canonical","Non-Canonical"))

### By Age
n.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))
n.non.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Non-Canonical") %>% summarise(n.non.can = n()))
n.tot <- merge(n.can, n.non.can)
n.tot$ratio.can <- n.tot$n.can/(n.tot$n.can + n.tot$n.non.can)
n.tot$ratio.non.can <- n.tot$n.non.can/(n.tot$n.can + n.tot$n.non.can)

### CANONICAL
df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.can, Cat = "Canonical"))
ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
  geom_bar(position="dodge", stat="identity")

### NONCANONICAL
df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.non.can, Cat = "Non-Canonical"))
ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
  geom_bar(position="dodge", stat="identity")

###BOTH

df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.can, Cat = "Canonical"),data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.non.can, Cat = "Non-canonical"))

ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
  geom_bar(position="dodge", stat="identity")
```

```{r ChildID}
### By Recording (=Child ID)
n.can <- as.data.frame(df.sub %>% group_by(ChildID) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))
n.non.can <- as.data.frame(df.sub %>% group_by(ChildID) %>% filter(Zoon_classif == "Non-Canonical") %>% summarise(n.non.can = n()))
n.tot <- merge(n.can, n.non.can)
n.tot$ratio.can <- n.tot$n.can/(n.tot$n.can + n.tot$n.non.can)
n.tot$ratio.non.can <- n.tot$n.non.can/(n.tot$n.can + n.tot$n.non.can)
df.plot <- rbind(data.frame(ChildID = as.factor(n.tot$ChildID), Ratio = n.tot$ratio.can, Cat = "Canonical"),
                 data.frame(ChildID = as.factor(n.tot$ChildID), Ratio = n.tot$ratio.non.can, Cat = "Non-Canonical"))

ggplot(df.plot, aes(fill=Cat, y=Ratio, x=ChildID)) + 
  geom_bar(position="dodge", stat="identity")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

---
title: "SLT_paper"
author: "AC"
date: "2020-03-06 (substantive version), latest minor edits 2020-07-17"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
library(scales)

@@

## History:

- 2020-08-05 final first version

## Read data in

```{r}
# read dataset composed with python
data_all <- read.csv("final_classifications_PU_zoon.csv")

@@

#relevel the factor so that it's easier to read
data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",
                                                             "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))

# create lab column with easier to read correspondance
data_all$lab<-as.character(data_all$Major_Choice)
data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
```

## Correspondence between lab & zooniverse annotation at the level of segments

Here we look at to what extent zooniverse and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").

@@

mycf

@@

### Precision

Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?

@@

### Recall

Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?

@@

## Child level descriptors

Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development. This is what we look at in this section.

```{r prepare data by child}

@@

#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
ztab=table(data_all$ChildID,data_all$Zoon_classif)
z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])

ltab=table(data_all$ChildID,data_all$lab)
l_lr=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
l_cr=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])

#put all the ratios together
if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) ratios=cbind(rownames(ztab),z_lr,z_cr,l_lr,l_cr) else print("oops this code needs to be more complex because we don't have the same kids for the two ratios")
colnames(ratios)[1]<-"ChildID"

#add age
ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean

merge(ratios,ages,by.x="ChildID",by.y="Group.1")->ratios
colnames(ratios)[dim(ratios)[2]]<-"Age"

#cbinding results in text, so we numerize the ratios
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) ratios[,thisvar]=as.numeric(as.character(ratios[,thisvar]))
summary(ratios)
```

We first look generally at two measures that have been found to relate to age:

- linguistic ratio = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e. we remove junk)
- canonical ratio = "Canonical"/("Canonical"+"Non-Canonical") (i.e. we remove junk + non-linguistic vocalizations)

As expected, linguistic ratio goes up with age.

Surprisingly, canonical ratio goes DOWN with age.

```{r corage}
prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
  myr=round(cor.test(ratios[,thisvar],ratios$Age)$estimate,3)
  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(ratios[,thisvar]~ratios$Age))
  }

@@

But the key thing for us: Are Zooniverse annotations describing children similar to lab annotations? The answer is clearly yes.

```{r corlab-zoo}

#Ling ratio
pdf("ling_rat_z_vs_l.pdf",height=5,width=5)
lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims)
  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()

  #CR
pdf("can_rat_z_vs_l.pdf",height=5,width=5)
lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims)
  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```
